The Concurrent Calculi Formalisation Benchmark

### 2024-02-21 short-paper/short-paper.tex
two small corrections to intro

In this spirit, we introduce a new collection of benchmark problems
specifically crafted to tackle common challenges encountered during
the mechanisation of formal models for concurrent and distributed
programming languages.
We particularly focus on process calculi.\todo{Why? - they are PL-like but also formal models in themselves}
Concurrent calculi are notably subtle: For instance, it took nine
years before the incorrect proof of type safety in the seminal paper
by Honda et al.~\cite{Honda1998} was rectified in Yoshida \&
Vasconcelos’s subsequent work\cite{Yoshida2007}.
Similarly, several key results in papers on session types have turned out
to be false later on, illustrating the
value of machine-checked proofs~\cite{Gay2020,10.1145/3290343}.

While results about concurrent
formalisms have already been mechanised (as we will discuss further
below), our experience is that choosing appropriate mechanisation
techniques and tools remains a significant challenge and that the
trade-offs are not well understood.  This often leads researchers
toward a trial-and-error approach, resulting in sub-optimal solutions,
wasted mechanisation effort, and techniques and results that are hard
to reuse. For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
mechanised proofs closely resemble the informal ones, while the
main challenge lies in getting the foundations right.

Our benchmark problems (detailed in \cref{app:challenges} and on our website) consider
\emph{in isolation} three key aspects that frequently pose challenges
when mechanising concurrency theory: \emph{linearity}, \emph{scope
  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
often address several of these aspects; however, we
see the combination of techniques as a next
step, as discussed in \cref{sec:going-beyond}.
Our benchmark challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.

>>>>

In this spirit, we introduce a new collection of benchmark problems
specifically crafted to tackle common challenges encountered during
the mechanisation of formal models for concurrent and distributed
programming languages.
We particularly focus on process calculi.\todo{Why? - they are PL-like but also formal models in themselves}
Concurrent calculi are notably subtle: for instance, it took nine
years before the incorrect proof of type safety in the seminal paper
by Honda et al.~\cite{Honda1998} was rectified in Yoshida \&
Vasconcelos’s subsequent work\cite{Yoshida2007}.
Similarly, several key results in papers on session types have turned out
to be false later on, illustrating the
value of machine-checked proofs~\cite{Gay2020,10.1145/3290343}.

While results about concurrent
formalisms have already been mechanised (as we will discuss further
below), our experience is that choosing appropriate mechanisation
techniques and tools remains a significant challenge and that the
trade-offs are not well understood.  This often leads researchers
toward a trial-and-error approach, resulting in sub-optimal solutions,
wasted mechanisation effort, and techniques and results that are hard
to reuse. For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
mechanised proofs closely resemble the informal ones, while the
main challenge lies in getting the infra-structure right.

Our benchmark problems (detailed in \cref{app:challenges} and on our website) consider
\emph{in isolation} three key aspects that frequently pose challenges
when mechanising concurrency theory: \emph{linearity}, \emph{scope
  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
often address several of these aspects at the same time; however, we
see the combination of techniques as a next
step, as discussed in \cref{sec:going-beyond}.
Our benchmark challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.
